---
title: "Live Cell Imaging Analysis"
output: html_document
date: "2023-11-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(ggExtra)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(gghalves)
library(cowplot)
library(gplots)
library(lattice)
library(ggthemes)
library(scales)
library(seqinr)
library(tidyverse)
library(Dict)
library(zoo)
library(emdbook)
library(PupillometryR)
library(lemon)
library(embed)
library(fpc)
library(dbscan)
library(viridis)
library(latex2exp)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(gtable)
library(readxl)
library(drc)
library(modelr)
```

## Establoshing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1),
  panel.spacing.y = unit(0.1, "lines"),
  panel.spacing.x = unit(0.1, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.35),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2)),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2)),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2)),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2)),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2)),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2)),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 5, margin = margin(r = 1)),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000", "#ccdba2", "#889466", "#c6b2d1", "#654875")
```

## Establishing paths to the data files and the folder to store results
```{r}
path_data <- "/path/to/LC_mitotic_defects.xlsx"
path_data_euclidean <- "/path/to/IF_C-604_euclidean_distance.csv"
path_results <- "/path/to/results/"
```

## Importing IF C-604 response data
```{r}
data_euclidean_distance <- read_csv(path_data_euclidean, show_col_types = FALSE)
data_euclidean_distance_summary <- reframe(data_euclidean_distance, dist_mean = mean(dist), dist_sd = sd(dist), .by = c(cell_line, treatment))

head(data_euclidean_distance_summary)
```

## Importing data
```{r}
tmp_sheets <- excel_sheets(path_data)
data_raw <- data.frame()

for (s in tmp_sheets) {
  data_raw <- bind_rows(data_raw, read_excel(path_data, sheet = s))
}

data_raw
```

## Processing data
```{r}
data_failed_cytokinesis <- data_raw %>%
  reframe(n = n(), .by = c(experiment, cell_line, C604, outcome)) %>%
  spread(key = outcome, value = n) %>%
  replace(is.na(.), 0) %>%
  mutate(prop = defect / (defect + normal))

data_failed_cytokinesis_summary <- data_failed_cytokinesis %>%
  reframe(prop_mean = mean(prop),
          prop_sd = sd(prop),
          .by = c(cell_line, C604))

data_failed_cytokinesis
data_failed_cytokinesis_summary
```

## Plotting proportions of defective mitotic events
```{r}
list_cell_lines <- c("HCC1395", "U2OS", "MM231", "RPE-1", "HeLa", "BT-549")

data_graph <- data_failed_cytokinesis_summary %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines),
         C604 = factor(C604, levels = c(0, 0.25, 0.5, 1, 2)))

data_graph_points <- data_failed_cytokinesis %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines),
         C604 = factor(C604, levels = c(0, 0.25, 0.5, 1, 2)))

graph_mitotic_defects <- ggplot() +
  
  geom_bar(data = data_graph, aes(x = C604, y = prop_mean),
           stat = "identity", position = "identity",
           colour = colour_pallette[6], fill = colour_pallette[9],
           width = 0.7, linewidth = 0.25) +
  
  geom_errorbar(data = data_graph, aes(x = C604, ymin = prop_mean, ymax = prop_mean + prop_sd),
                linewidth = 0.25, width = 0.5) +
  
  geom_point(data = data_graph_points, aes(x = C604, y = prop, group = experiment),
           colour = colour_pallette[6], fill = "#FFFFFF",
           size = 0.85, stroke = 0.35, shape = 21,
           position = position_dodge(0.75)) +
  
  facet_grid(. ~ cell_line) +
  labs(x = "C-604 (µM)", y = "Defective\nmitosis") +
  scale_y_continuous(expand = c(0, 0),limits = c(0, 1.1), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = scales::percent_format(accuracy = 1)) +
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, vjust = 0.5, hjust = 0),
        panel.grid.major.y = element_line(colour = "#D4D3CF", linewidth = 0.15),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#D4D3CF", linewidth = 0.15),
        panel.grid.minor.x = element_blank())
  
ggsave(paste(path_results, "graph_defective_mitosis.pdf", sep = ""), graph_mitotic_defects,
       width = 72.5, height = 22.5, units = "mm", device = cairo_pdf)
```

## Plotting mitotic timing
```{r}
list_cell_lines <- c("HCC1395", "U2OS", "MM231", "RPE-1", "HeLa", "BT-549")

data_graph <- data_raw %>%
  reframe(timing_median = median(timing),
          .by = c(experiment, cell_line, C604)) %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines),
         C604 = factor(C604, levels = c(0, 0.25, 0.5, 1, 2)))

data_graph_points <- data_raw %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines),
         C604 = factor(C604, levels = c(0, 0.25, 0.5, 1, 2)))

data_graph_p <- data_graph %>%
  group_by(cell_line) %>%
  group_split() %>%
  lapply((function (data) {
    
    tmp.control <- filter(data, C604 == 0)
    tmp.data <- filter(data, C604 != 0) %>%
      group_by(C604) %>%
      group_split() %>%
      lapply((function (data_2) {
        
        tmp.p <- t.test(tmp.control$timing_median, data_2$timing_median, var.equal = TRUE)[3][[1]]
        tmp.output <- data.frame(cell_line = unique(data_2$cell_line),
                                 C604 = unique(data_2$C604),
                                 p = tmp.p)
        return(tmp.output)
        print(tmp.output)
        }))
    
  })) %>%
  bind_rows() %>%
  mutate(p_adj = p.adjust(p, method = "BH")) %>%
  mutate(label = case_when(p_adj >= 0.05 ~ "ns",
                           p_adj < 0.05 & p_adj >= 0.01 ~ "*",
                           p_adj < 0.01 & p_adj >= 0.001 ~ "**",
                           p_adj < 0.001 ~ "***")) %>%
  filter(p_adj < 0.05)

graph_mitotic_timing <- ggplot() +

  geom_quasirandom(data = data_graph_points, aes(x = C604, y = timing, group = experiment),
                   colour = colour_pallette[9],
                   size = 0.5, shape = 16, alpha = 1) +
  
  geom_point(data = data_graph, aes(x = C604, y = timing_median, group = experiment),
           colour = colour_pallette[6], fill = "#FFFFFF",
           size = 0.85, stroke = 0.3, shape = 21,
           position = position_dodge(0.75)) +
  
  geom_text(data = data_graph_p, aes(x = C604, y = 1750, label = label), size = 6/.pt, family = "Arial", colour = colour_pallette[10]) +
  
  facet_grid(. ~ cell_line) +
  labs(x = "C-604 (µM)", y = "Mitotic\ntiming (min)") +
  scale_y_continuous(expand = c(0, 0), trans = "log2", breaks = c(20, 80, 320, 1280), limits = c(12, 4000)) +
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, vjust = 0.5, hjust = 0),
        panel.grid.major.y = element_line(colour = "#D4D3CF", linewidth = 0.15),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_line(colour = "#D4D3CF", linewidth = 0.15),
        panel.grid.minor.x = element_blank())
  
ggsave(paste(path_results, "graph_mitotic_timing.pdf", sep = ""), graph_mitotic_timing,
       width = 72.5, height = 22.5, units = "mm", device = cairo_pdf)
```

## Plotting a correlation scatterplot of frequencies of mitotic exit defects and long-term cellular responses to C-604 (Euclidean distances)
```{r}
data_graph <- data_failed_cytokinesis_summary %>%
  rename(treatment = C604) %>%
  merge(data_euclidean_distance_summary, on = c(cell_line, treatment)) %>%
  mutate(cell_line = factor(cell_line, levels = c("HCC1395", "U2OS", "MM231", "RPE-1", "HeLa", "BT-549")))

data_corr <- data_graph %>%
  group_by(cell_line) %>%
  group_split() %>%
  lapply((function (data) {

    tmp.pearson <- cor.test(data$prop_mean, data$dist_mean, method = "pearson")
    tmp.lm <- lm(data$prop_mean ~ data$dist_mean)
    tmp.output <- data.frame(cell_line = unique(data$cell_line),
                             pearson = round(tmp.pearson$estimate, 3),
                             pearson_p = tmp.pearson$p.value,
                             lm_R_squared = summary(tmp.lm)$r.squared)
    
    return(tmp.output)
    
  })) %>%
  bind_rows()


pearson <- cor.test(data_graph$prop_mean, data_graph$dist_mean, method = "pearson")

graph_corr_distance_defects <- ggplot(data = data_graph) +
  
  geom_point(aes(x = dist_mean, y = prop_mean), size = 2, stroke = 0.25, shape = 21, colour = "#FFFFFF", fill = colour_pallette[10]) +
  geom_text(data = data_corr, aes(x = 90, y = 0.05, label = pearson), family = "Arial", size = 6/.pt, hjust = 1, colour = colour_pallette[10]) +
  scale_x_continuous(expand = c(0.05, 0.05), limits = c(0, 90), breaks = c(0, 20, 40, 60, 80)) +
  scale_y_continuous(expand = c(0.05, 0.05),limits = c(0, 1.05), breaks = c(0, 0.25, 0.5, 0.75, 1), labels = scales::percent_format(accuracy = 1)) +
  labs(x = "Euclidean distance of cellular responses to C-604", y = "Defective\nmitotic exit") +
  
  facet_grid(. ~ cell_line) +
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"))
  
ggsave(paste(path_results, "graph_corr_distance_defects.pdf", sep = ""), graph_corr_distance_defects,
       width = 80, height = 22.5, units = "mm", device = cairo_pdf)
```
