---
title: "Cell line panel characterisation (WB)"
output: html_document
date: "2024-06-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(ggExtra)
library(ggtext)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(gghalves)
library(ggrepel)
library(ggvenn)
library(cowplot)
library(gplots)
library(lattice)
library(ggthemes)
library(scales)
library(seqinr)
library(tidyverse)
library(Dict)
library(zoo)
library(emdbook)
library(PupillometryR)
library(lemon)
library(embed)
library(fpc)
library(dbscan)
library(viridis)
library(latex2exp)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(readxl)
library(qvalue)
library(tidymodels)
library(embed)
library(fpc)
library(dbscan)
library(corrr)
library(FactoMineR)
library(factoextra)
```

## Establishing paths to the data files and the folder to store results
```{r}
path_data <- "/path/to/WB_B55a-GWL-PP2C-ENSA_protein-levels.xlsx"
path_data_ED50 <- "/path/to/CF_C-604_ED50.csv"
path_results <- "/path/to/results/"
```

## Establishing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1, family = "Arial"),
  panel.spacing.y = unit(0.1, "lines"),
  panel.spacing.x = unit(0.1, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.5),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2), vjust = 0, family = "Arial"),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), family = "Arial"),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 3), family = "Arial"),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), family = "Arial"),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 6, margin = margin(r = 1), family = "Arial"),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000", "#ccdba2", "#889466", "#c6b2d1", "#654875")
```

## Importing experimental ED50 values of C-604
```{r}
data_ED50 <- read_csv(path_data_ED50, show_col_types = FALSE)

data_ED50
```

## Importing and processing protein expression (WB) data
```{r}
data_expression <- read_xlsx(path_data) %>%
  gather(key = gene, value = expression, 3:ncol(.)) %>%
  mutate(gene = paste(gene, "_raw", sep = "")) %>%
  spread(key = gene, value = expression) %>%
  mutate(PPP2R2A_normalised = PPP2R2A_raw / GAPDH_raw,
         MASTL_normalised = MASTL_raw / GAPDH_raw,
         ENSA_normalised = ENSA_raw / GAPDH_raw,
         PPP2C_normalised = PPP2C_raw / GAPDH_raw) %>%
  gather(key = gene, value = expression, 3:ncol(.)) %>%
  separate(gene, into = c("gene", "data"))

data_expression_scaled <- data_expression %>%
  filter(data == "normalised") %>%
  group_by(experiment, gene) %>%
  mutate(expression = expression / max(expression)) %>%
  ungroup() %>%
  merge(data_ED50, on = cell_line) %>%
  mutate(cell_line = factor(cell_line, levels = unique(cell_line))) 

data_expression_scaled_summary <- data_expression_scaled %>%
  reframe(expression_mean = mean(expression),
          expression_sd = sd(expression),
          .by = c(cell_line, gene)) %>%
  merge(data_ED50, on = cell_line)

data_expression
```

## Predicting ED50 values using mathematical models
```{r}
model1 <- function(PPP2R2A, MASTL) {return(0.399 * MASTL / PPP2R2A + 0.406)}

data_ED50_prediction <- data_expression_scaled %>%
  dplyr::select(experiment, cell_line, gene, expression) %>%
  spread(key = gene, value = expression) %>%
  mutate(ED50_prediction = model1(PPP2R2A, MASTL))

data_ED50_prediction_summary <- data_ED50_prediction %>%
  reframe(ED50_prediction_mean = mean(ED50_prediction), ED50_prediction_sd = sd(ED50_prediction),
          .by = cell_line) %>%
  merge(data_ED50, on = cell_line) %>%
  dplyr::select(group, cell_line, ED50_mean, ED50_sd, ED50_prediction_mean, ED50_prediction_sd)

data_ED50_prediction_summary
```

## Calculating correlation coefficients
```{r}
data_corr <- data_expression_scaled_summary %>%
  group_by(gene) %>%
  group_split() %>%
  lapply((function (data) {
    
    tmp.expression = data$expression_mean
    tmp.ED50 = data$ED50_mean
    tmp.corr <- cor.test(tmp.ED50, tmp.expression, method = "pearson")
    tmp.lm <- lm(tmp.ED50 ~ tmp.expression)
    
    data_output <- data.frame(
      gene = unique(data$gene),
      pearson_corr = tmp.corr$estimate,
      pearson_p = tmp.corr$p.value,
      lm_slope = summary(tmp.lm)$coef[2],
      lm_R_squared = summary(tmp.lm)$r.squared,
      lm_R_squared_adj = summary(tmp.lm)$adj.r.squared)
    
    return(data_output)

  })) %>%
  bind_rows()

data_corr
```

## Plotting expression levels
```{r}
list_cell_lines <- c("HCC1143", "HCC1395", "HCC1937", "NCI-H226", "NCI-H2228", "769-P",
                      "U2OS", "MM231", "RPE-1", "HT-1080", "HeLa", "BT-549")

data_graph <- data_expression_scaled_summary %>%
  mutate(gene = factor(gene, levels = c("MASTL", "PPP2R2A", "PPP2C", "ENSA"), labels = c("\nGWL", "\nB55α", "\nPPP2Cα/β", "ENSA\nARPP19")),
         cell_line = factor(cell_line, levels = list_cell_lines))

data_graph_points <- data_expression_scaled %>%
  mutate(gene = factor(gene, levels = c("MASTL", "PPP2R2A", "PPP2C", "ENSA"), labels = c("\nGWL", "\nB55α", "\nPPP2Cα/β", "ENSA\nARPP19")),
         cell_line = factor(cell_line, levels = list_cell_lines))

data_graph_corr <- data_corr %>%
  mutate(gene = factor(gene, levels = c("MASTL", "PPP2R2A", "PPP2C", "ENSA"), labels = c("\nGWL", "\nB55α", "\nPPP2Cα/β", "ENSA\nARPP19")))

graph_expression <- ggplot(data = data_graph,
                           aes(x = cell_line, y = expression_mean)) +
  
  geom_bar(aes(x = cell_line, y = expression_mean, fill = gene),
           stat = "identity", position = "identity", colour = colour_pallette[6],
           linewidth = 0.25, width = 0.6) +
  
  geom_errorbar(aes(x = cell_line,
                    ymin = expression_mean,
                    ymax = expression_mean + expression_sd),
                width = 0.5, linewidth = 0.25, colour = colour_pallette[6]) +
  
  geom_point(data = data_graph_points,
             aes(x = cell_line, y = expression, group = experiment),
             position = position_dodge(0.75),
             shape = 21, size = 0.85, stroke = 0.35,
             fill = "#FFFFFF") +
  
  geom_text(data = filter(data_graph_corr, gene %in% c("\nB55α", "\nPPP2Cα/β")), aes(x = 0.5, y = 0.95, label = paste("r = ", round(pearson_corr, 2), sep = "")),
            size = 2.11, vjust = 0, hjust = -0.1) +
  
  geom_text(data = filter(data_graph_corr, gene %in% c("ENSA\nARPP19", "\nGWL")), aes(x = 11.5, y = 0.95, label = paste("r = ", round(pearson_corr, 2), sep = "")),
            size = 2.11, vjust = 0, hjust = 0.8) +

  scale_x_discrete(expand = c(0.05, 0.05)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1.2), breaks = c(0, 0.25, 0.5, 0.75, 1)) +
  scale_fill_manual(values = alpha(c(colour_pallette[5], colour_pallette[2], colour_pallette[4], colour_pallette[1]), 0.5)) +
  labs(x = "", y = "Expression (normalised, scaled)") +
  facet_grid(gene ~ ., scales = "free_y") +
  
  graph_theme +
  theme(
    axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"),
    plot.margin = margin(b = 0.2, l = 0.3, t = -0.575, r = 0.2, unit = "lines"),
    axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 5)),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank())

## Plotting ED50 values

list_cell_lines <- c("HCC1143", "HCC1395", "HCC1937", "NCI-H226", "NCI-H2228", "769-P",
                      "U2OS", "MM231", "RPE-1", "HT-1080", "HeLa", "BT-549")

data_graph <- data_expression_scaled %>%
  dplyr::select(cell_line, ED50_mean, ED50_sd) %>%
  unique() %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines))

graph_ED50 <- ggplot(data = data_graph,
                     aes(x = cell_line, y = ED50_mean)) +
  
  geom_bar(aes(x = cell_line, y = ED50_mean),
           stat = "identity", position = "identity",
           colour = colour_pallette[6], fill = colour_pallette[6],
           linewidth = 0.25, width = 0.6) +
  
  geom_errorbar(aes(x = cell_line,
                    ymin = ED50_mean,
                    ymax = ED50_mean + ED50_sd),
                width = 0.5, linewidth = 0.25, colour = colour_pallette[6]) +

  scale_x_discrete(expand = c(0.05, 0.05)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 2.8), breaks = c(0, 0.8, 1.6, 2.4)) +
  labs(x = "", y = bquote(ED[50]~"(µM)")) +
  
  graph_theme +
  theme(
    panel.border = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.line = element_line(linewidth = 0.25, colour = "#000000"),
    plot.margin = margin(b = 0, l = 0.3, t = 0.3, r = 0.2, unit = "lines"),
    axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 5)),
    panel.grid.major.y = element_blank(),
    panel.grid.major.x = element_blank())

## Creating a composite graph of expresion and ED50 values

graph_composite <- plot_grid(graph_ED50,
                             graph_expression,
                             ncol = 1, nrow = 2,
                             rel_widths = c(1, 1),
                             rel_heights = c(1, 5),
                             align = "hv",
                             axis = "lr")

ggsave(paste(path_results, "graph_composite.pdf", sep = ""), graph_composite,
       width = 50, height = 69.085 , units = "mm", device = cairo_pdf)
```
## Comparing empiric and modeled data
```{r}
list_cell_lines <- c("HCC1143", "HCC1395", "HCC1937", "NCI-H226", "NCI-H2228", "769-P",
                      "U2OS", "MM231", "RPE-1", "HT-1080", "HeLa", "BT-549")

data_graph <- data_ED50_prediction_summary %>%
  mutate(cell_line = factor(cell_line, levels = list_cell_lines))

pearson_cor <- cor.test(data_graph$ED50_mean, data_graph$ED50_prediction_mean, method = "pearson")
linear_model <- lm(data_graph$ED50_prediction_mean~data_graph$ED50_mean)
lm_R_squared = round(summary(linear_model)$r.squared, 2)

graph_graph_model_vs_empiric <- ggplot(data = data_graph, aes(x = ED50_mean, y = ED50_prediction_mean)) +
  
  geom_smooth(method = "lm", formula = y~x,
              colour = colour_pallette[10], fill = colour_pallette[9],
              linewidth = 0.4) +
  
  geom_point(size = 1.75, stroke = 0.5, shape = 21, colour = colour_pallette[10], fill = "#FFFFFF") +
  
  annotate("text", x = 0.2, y = 2, label = paste("r = ", round(pearson_cor$estimate, 2), sep = ""),
           family = "Arial", size = 6/.pt, hjust = 0, vjust = 0.2, colour = colour_pallette[6]) +
  
  annotate("text", x = 0.2, y = 1.7, label = bquote(R^2~"="~.(lm_R_squared)),
           family = "Arial", size = 6/.pt, hjust = 0, vjust = 0.2, colour = colour_pallette[6]) +
  
  labs(x = bquote(Experimental~ED[50]~(µM)), y = bquote("Model *ED"[50]~(µM))) +
  scale_x_continuous(expand = c(0, 0), breaks = c(0.1, 0.4, 0.8, 1.2, 1.6, 2)) +
  scale_y_continuous(expand = c(0, 0), breaks = c(0.1, 0.4, 0.8, 1.2, 1.6, 2)) +
  coord_cartesian(clip = 'off') +
  graph_theme +
  theme(panel.border = element_blank(),
        axis.line = element_line(colour = "#000000", linewidth = 0.25))

ggsave(paste(path_results, "graph_model_vs_empiric.pdf", sep = ""), graph_graph_model_vs_empiric,
       width = 37.5, height = 35 , units = "mm", device = cairo_pdf)
```
