---
title: "GLW inhibition (off-target kinases siRNA testing)"
output: html_document
date: "2023-11-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(ggExtra)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(gghalves)
library(cowplot)
library(gplots)
library(lattice)
library(ggthemes)
library(scales)
library(tidyverse)
library(viridis)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(gtable)
library(embed)
library(fpc)
library(dbscan)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(corrr)
library(transport)
```

## Establishing paths to the data file and the folder to store results
```{r}
path_data <- "/path/to/IF_single-cell_off-target_siRNAs.csv"
path_results <- "/path/to/results/"
```

## Establoshing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1, family = "Arial"),
  panel.spacing.y = unit(0.1, "lines"),
  panel.spacing.x = unit(0.1, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.5),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2), vjust = 0),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2)),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 3)),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2)),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2)),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2)),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 6, margin = margin(r = 1)),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000")
```

## Establishing a function to normalise the raw IF data
```{r}
fun_normalise <- function (data, values, ref, bins = 100) {

  # Constructing the subset data-set containing the control data
  
  for (v in values) {
    
    if (v == "DAPI_total") {data.ref <- filter(data, condition %in% ref, EdU == "negative")}
    else {data.ref <- subset(data, condition %in% ref)}

    # log10-transformation of the data
    data.ref <- mutate(data.ref, value_log2 = log2(!!sym(v)))
    # Construction of the histogram
    histogram <- hist(x = data.ref$value_log2,
                      breaks = bins,
                      main = paste(unique(data$experiment), unique(data$cell_line), sep = " "),
                      xlab = paste("log2", v, sep = " "))
    
    # Selection of the histogram index (bin) with the highest number of observations
    index_max <- match(max(histogram$counts), histogram$counts)
    
    # Determination of the mid-log10-value of the bin with the highest number of observations
    value_max <- histogram$mids[index_max]
    
    # Determination of the normalisation factor (reversing the log10-transformation)
    normalisation_factor <- 2 ^ value_max
    
    # Normalisation of the original values
    data <- mutate(data, !!sym(paste(v, "norm", sep = "_")) := !!sym(v) / normalisation_factor)
    
    # Normalised DAPI intensities are multiplied by 2 so G1 populations center around 2 (2N) and G2 populations center around 4 (4N)
    if (v == "DAPI_total") {data <- mutate(data, !!sym(paste(v, "norm", sep = "_")) := 2 * !!sym(paste(v, "norm", sep = "_")))}
  }
  return(data)
}
```

## Establishing a function to establish density of points on a scatter plot
```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
```

## Creating a dataframe containing Hoechst, EdU and p21 intensity thresholds
```{r}
data_thresholds <- bind_rows(data.frame(feature = "Hoechst", T1 = 1.5, T2 = 3, T3 = 6),
                             data.frame(feature = "EdU", T1 = 1.5, T2 = NA, T3 = NA), 
                             data.frame(feature = "p21", T1 = 1.3, T2 = NA, T3 = NA))

data_thresholds
```

## Importing data
```{r}
data_raw <- read.csv(path_data, sep = ",")

head(data_raw)
```

## Normalising data
```{r}
data_normalised = data_raw %>%
  mutate(experiment = unlist(str_extract_all(data_raw$experiment, "\\d+"))) %>%
  
  reframe(nuclei_count = n(),
          nuclei_area = sum(area_nucleus),
          DAPI_total = sum(integrated_int_DAPI),
          EdU_mean = mean(intensity_mean_EdU_nucleus) / mean(intensity_mean_EdU_cyto),
          p21_mean = mean(intensity_mean_p21_nucleus),
          .by = c(experiment, plate_id, well, well_id, image_id, cell_line, condition, treatment, siRNA, cyto_id, area_cell)) %>%
  
  group_by(experiment) %>%
  group_split() %>%
  lapply(fun_normalise, values = c("EdU_mean", "p21_mean"), ref = c("CTR_0")) %>%
  bind_rows() %>%
  mutate(EdU = case_when(EdU_mean_norm <= filter(data_thresholds, feature == "EdU")$T1 ~ "negative",
                         EdU_mean_norm > filter(data_thresholds, feature == "EdU")$T1 ~ "positive"),
         p21 = case_when(p21_mean_norm <= filter(data_thresholds, feature == "p21")$T1 ~ "negative",
                         p21_mean_norm > filter(data_thresholds, feature == "p21")$T1 ~ "positive")) %>%
  group_by(experiment, cell_line) %>%
  group_split() %>%
  lapply(fun_normalise, values = c("DAPI_total", "area_cell", "nuclei_area"), ref = c("CTR_0")) %>%
  bind_rows() %>%
  
  mutate(Hoechst = case_when(DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T1 ~ "Debris",
                             DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T2 ~ "2N",
                             DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T3 ~ "4N",
                             DAPI_total_norm > filter(data_thresholds, feature == "Hoechst")$T3 ~ "8N+")) %>%
  
  mutate(cell_cycle = case_when(Hoechst == "Debris" ~ "Debris",
                                Hoechst == "2N" & EdU == "negative" ~ "2N",
                                Hoechst == "4N" & EdU == "negative" ~ "4N",
                                Hoechst == "2N" & EdU == "positive" ~ "S",
                                Hoechst == "4N" & EdU == "positive" ~ "S",
                                Hoechst == "8N+" ~ "8N+",
                                .default = "Unassigned"))

head(data_normalised)
```

## Establishing cell counts
```{r}
data_cell_counts <- data_normalised %>%
  filter(cell_cycle != "Debris") %>%
  reframe(count = n(), .by = c(experiment, cell_line, siRNA, treatment))
  
print(data_cell_counts)
```

## Creating random subset data
```{r}
set.seed(123)
data_sample <- data_normalised %>%
  group_by(experiment, cell_line, siRNA, treatment) %>%
  sample_n(500) %>%
  ungroup()

head(data_sample)
```

## Constructing the input dataset for the Uniform Manifold Approximation and Projection (UMAP) fitting
```{r}
data_umap_input <- data_sample %>%
  
  mutate(DAPI_log2 = log2(DAPI_total_norm),
         EdU_log2 = log2(EdU_mean_norm),
         p21_log2 = log2(p21_mean_norm),
         area_cell_log2 = log2(area_cell),
         nuclei_count) %>%
  
  dplyr::select(experiment, cell_line, siRNA, treatment, DAPI_log2, EdU_log2, p21_log2, area_cell_log2, nuclei_count)

head(data_umap_input)
```

## Fitting the Uniform Manifold Approximation and Projection (UMAP)
```{r}
set.seed(123)
data_umap <- recipe(experiment + cell_line + siRNA + treatment ~ ., data = data_umap_input) %>% 
  step_normalize(all_predictors()) %>% 
  step_umap(all_predictors(), num_comp = 2) %>% 
  prep() %>% 
  juice()

head(data_umap)
```

## Computing Earth Mover's (Wasserstein) Distance
```{r}
list_siRNAs <- c("CTR", "MASTL", "NUAK1", "MELK", "ULK2", "MAP3K9", "HIPK2", "CHEK1")

data_umap_wasserstein <- data_umap %>%
  filter(siRNA %in% list_siRNAs) %>%
  group_by(cell_line) %>%
  group_split() %>%
  lapply((function (data) {
    
    tmp.data <- data %>%
      group_by(siRNA, treatment) %>%
      group_split()
    
    tmp.output <- data.frame()
    
    for (df.a in tmp.data) {
      
      tmp.dataset.a <- paste(unique(df.a$siRNA), unique(df.a$treatment), sep = "_")
      umap.a <- matrix(c(df.a$UMAP1, df.a$UMAP2), ncol = 2)
      
      for (df.b in tmp.data) {
        
        tmp.dataset.b <- paste(unique(df.b$siRNA), unique(df.b$treatment), sep = "_")
        umap.b <- matrix(c(df.b$UMAP1, df.b$UMAP2), ncol = 2)
        
        # Compute pairwise cost matrix (Euclidean distances)
        cost.matrix <- as.matrix(dist(rbind(umap.a, umap.b)))
        cost.matrix <- cost.matrix[1:nrow(umap.a), (nrow(umap.a) + 1):nrow(cost.matrix)]
        
        # Assign uniform weights since points are empirical distributions
        n1 <- nrow(umap.a)
        n2 <- nrow(umap.b)
        weights1 <- rep(1/n1, n1)  # Normalize weights to sum to 1
        weights2 <- rep(1/n2, n2)
        
        # Compute Earth Mover's Distance
        emd.value <- wasserstein(weights1, weights2, costm = cost.matrix, p = 1)
        tmp.result <- data.frame(cell_line = unique(data$cell_line),
                                 dataset_a = tmp.dataset.a,
                                 dataset_b = tmp.dataset.b,
                                 emd = emd.value)
        
        tmp.output <- bind_rows(tmp.output, tmp.result)
      }
    }
    return(tmp.output)
  })) %>%
  bind_rows() %>%
  separate(col = dataset_a, into = c("siRNA_a", "treatment_a"), sep = "_", remove = FALSE) %>%
  separate(col = dataset_b, into = c("siRNA_b", "treatment_b"), sep = "_", remove = FALSE) %>%
  arrange(emd) %>%
  mutate(rank = 1:n())
```

## Statistical comparison of Earth Mover's (Wasserstein) Distances
```{r}
data_umap_wasserstein_subset <- data_umap_wasserstein %>%
  filter(siRNA_a == "CTR", treatment_a == "2") %>%
  dplyr::rename(siRNA = siRNA_b, treatment = treatment_b)

for (s in unique(data_umap_wasserstein_subset$siRNA)[!unique(data_umap_wasserstein_subset$siRNA) %in% c("CTR", "MASTL")]) {
  
  tmp.GWL <- filter(data_umap_wasserstein_subset, siRNA == "MASTL")$emd
  tmp.siRNA <- filter(data_umap_wasserstein_subset, siRNA == s)$emd
  print(s)
  print(wilcox.test(tmp.GWL, tmp.siRNA, alternative = "two.sided", paired = FALSE))
}

```

## Plotting Wasserstein distances from (siCTR, 2 µM C-604)
```{r}
list_siRNAs <- c("CTR", "MASTL", "NUAK1", "MELK", "ULK2", "MAP3K9", "HIPK2", "CHEK1")
data_graph_emd <- data_umap_wasserstein_subset %>%
  mutate(siRNA = factor(siRNA, levels = list_siRNAs, labels = c("CTR", "GWL", "NUAK1", "MELK", "ULK2", "MAP3K9", "HIPK2", "CHEK1")))
  

graph_emd <- ggplot(data = data_graph_emd, aes(x = treatment, y = emd, colour = cell_line, fill = cell_line, group = cell_line, shape = cell_line)) +
  
  geom_vline(xintercept = 1.5, linewidth = 0.25, colour = colour_pallette[1]) +
  geom_vline(xintercept = 2.5, linewidth = 0.25, colour = colour_pallette[1]) +
  geom_hline(yintercept = 2, linewidth = 0.25, colour = colour_pallette[1]) +
  
  geom_point(position = position_dodge(0.5), size = 1.75, stroke = 0.35, colour = "#FFFFFF", fill = "#FFFFFF") +
  geom_point(position = position_dodge(0.5), size = 1.75, stroke = 0.35, colour = "#000000") +

  facet_grid(. ~ siRNA) +
  scale_y_continuous(limits = c(0, 8), breaks = seq(0, 8)) +
  scale_colour_manual(values = c(colour_pallette[3], colour_pallette[1])) +
  scale_fill_manual(values = alpha(c(colour_pallette[2], colour_pallette[4]), 0.5)) +
  scale_shape_manual(values = c(21, 22)) +
  labs(x = "GWL inhibitor C-604 (µM)", y = "Wasserstein distance from\nsiCTR + C-604 (2 µM)") +
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, angle = -90, hjust = 0, vjust = 0.5, margin = margin(t = 2)))

ggsave(paste(path_results, "graph_emd.pdf", sep = ""), graph_emd,
       width = 72.5, height = 35, units = "mm", device = cairo_pdf)
```

## Plotting combined UMAP data
```{r}
list_siRNAs <- c("CTR", "MASTL", "NUAK1", "MELK", "ULK2", "MAP3K9", "HIPK2", "CHEK1")
data_graph_umap <- filter(data_umap, siRNA %in% list_siRNAs) %>%
  mutate(siRNA = factor(siRNA, levels = list_siRNAs, labels = c("CTR", "GWL", "NUAK1", "MELK", "ULK2", "MAP3K9", "HIPK2", "CHEK1"))) %>%
  group_by(cell_line, siRNA, treatment) %>%
  mutate(density = get_density(UMAP1, UMAP2, n = 500, h = c(10, 10))) %>%
  ungroup()

graph_umap_combined <- ggplot() +

  geom_point(data = data_graph_umap,
             aes(x = UMAP1, y = UMAP2, colour = density),
             size = 0.25, alpha = 0.2, shape = 16) +
  
  scale_x_continuous(expand = c(0.12, 0.12), limits = c(-15, 15), breaks = c(-12, 0, 12)) +
  scale_y_continuous(expand = c(0.12, 0.12), limits = c(-15, 15), breaks = c(-12, 0, 12)) +
  
  scale_colour_gradient2(low = colour_pallette[3], mid = colour_pallette[4], high = colour_pallette[2],
                       midpoint = quantile(data_graph_umap$density, 0.5),
                       limits = c(quantile(data_graph_umap$density, 0.1), quantile(data_graph_umap$density, 0.9)),
                       na.value = colour_pallette[1],
                       oob = scales::oob_squish) +
  
  facet_nested(siRNA ~ cell_line + treatment,
               nest_line = element_line(linetype = 1, linewidth = 0.25),
               resect = unit(0.04, "in"),
               strip = strip_nested(clip = "off")) +
  
  labs(colour = "Density (a.u.)") +
  
  graph_theme +
  theme(legend.position = "top",
        legend.justification = "center",
        legend.key = element_rect(colour = "#000000", linewidth = 1),
        legend.key.width = unit(20, 'pt'),
        legend.key.height = unit(4, 'pt'),
        legend.text = element_text(vjust = -0.5),
        legend.margin = margin(b = -9),
        legend.spacing.x = unit(1, "pt"),
        legend.title = element_text(colour = "#000000", size = 6, family = "Arial", margin = margin(b = 2))) +
  
  guides(colour = guide_colorbar(frame.colour = "#000000", frame.linewidth = 0.25,
                                 draw.ulim = FALSE, draw.llim = FALSE,
                                 ticks.colour = colour_pallette[7],
                                 title.position = "top", title.hjust = 0.5))

ggsave(paste(path_results, "graph_umap_combined.pdf", sep = ""), graph_umap_combined,
       width = 70, height = 90, units = "mm", device = cairo_pdf)
```
