---
title: "Treatment with C-604 and MKI-2"
output: html_document
date: "2024-01-04"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(data.table)
library(ggplot2)
library(ggforce)
library(ggExtra)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(gghalves)
library(cowplot)
library(gplots)
library(lattice)
library(ggthemes)
library(scales)
library(seqinr)
library(tidyverse)
library(Dict)
library(viridis)
library(latex2exp)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(gtable)
library(rstatix)
library(effsize)
```

## Establishing paths to the data file and the folder to store results
```{r}
path_data <- "/path/to/IF_single-cell_C-604_vs_MKI-2.csv"
path_results <- "/path/to/results/"
```

## Establoshing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1, family = "Arial"),
  panel.spacing.y = unit(0.05, "lines"),
  panel.spacing.x = unit(0.05, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.5),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2), family = "Arial"),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), family = "Arial"),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 3), family = "Arial"),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), family = "Arial"),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 6, margin = margin(r = 1), family = "Arial"),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000", "#ccdba2", "#889466")
```

## Establishing a function to normalise the raw IF data
```{r}
fun_normalise <- function (data, values, ref, bins = 100) {

  # Constructing the subset data-set containing the control data
  
  for (v in values) {
    
    if (v == "DAPI_total") {data.ref <- filter(data, condition %in% ref, EdU == "negative")}
    else {data.ref <- subset(data, condition %in% ref)}

    # log10-transformation of the data
    data.ref <- mutate(data.ref, value_log2 = log2(!!sym(v)))
    # Construction of the histogram
    histogram <- hist(x = data.ref$value_log2,
                      breaks = bins,
                      main = paste(unique(data$experiment), unique(data$cell_line), sep = " "),
                      xlab = paste("log2", v, sep = " "))
    
    # Selection of the histogram index (bin) with the highest number of observations
    index_max <- match(max(histogram$counts), histogram$counts)
    
    # Determination of the mid-log10-value of the bin with the highest number of observations
    value_max <- histogram$mids[index_max]
    
    # Determination of the normalisation factor (reversing the log10-transformation)
    normalisation_factor <- 2 ^ value_max
    
    # Normalisation of the original values
    data <- mutate(data, !!sym(paste(v, "norm", sep = "_")) := !!sym(v) / normalisation_factor)
    
    # Normalised DAPI intensities are multiplied by 2 so G1 populations center around 2 (2N) and G2 populations center around 4 (4N)
    if (v == "DAPI_total") {data <- mutate(data, !!sym(paste(v, "norm", sep = "_")) := 2 * !!sym(paste(v, "norm", sep = "_")))}
  }
  return(data)
}
```

## Establishing a function to establish density of points on a scatter plot
```{r}
get_density <- function(x, y, ...) {
  dens <- MASS::kde2d(x, y, ...)
  ix <- findInterval(x, dens$x)
  iy <- findInterval(y, dens$y)
  ii <- cbind(ix, iy)
  return(dens$z[ii])
}
```

## Creating a dataframe containing Hoechst, EdU and p21 intensity thresholds
```{r}
data_thresholds <- bind_rows(data.frame(feature = "Hoechst", T1 = 1.5, T2 = 3, T3 = 6),
                             data.frame(feature = "EdU", T1 = 1.5, T2 = NA, T3 = NA), 
                             data.frame(feature = "p21", T1 = 1.5, T2 = NA, T3 = NA))

data_thresholds
```

## Establishing a function to compute statistics
```{r}
fun_stats <- function (data, group_vars = c("cell_line", "C604", "RO3306"), val = "prop") {
  
  data_nested <- data %>%
    group_by(across(all_of(group_vars))) %>%
    nest()
  
  data_output <- data.frame()
  
  for (i in 1:nrow(data_nested)) {
    data_group_1 <- data_nested[i, ] %>%
      unite(col = "group", group_vars) %>%
      unnest(data)
    
    for (i in 1:nrow(data_nested)) {
      data_group_2 <- data_nested[i, ] %>%
        unite(col = "group", group_vars) %>%
        unnest(data)
      
      ## Normality test
      if (sd(data_group_1[[val]]) == 0){
        shapiro_1 <- NA
      } else {
        shapiro_1 <- shapiro.test(data_group_1[[val]])$p.value
      }
      if (sd(data_group_2[[val]]) == 0){
        shapiro_2 <- NA
      } else {
        shapiro_2 <- shapiro.test(data_group_2[[val]])$p.value
      }
      
      ## Cohen size effect
      cohen_d <- cohen.d(data_group_1[[val]], data_group_2[[val]])$estimate
      ## t-test
      t <- t.test(data_group_1[[val]], data_group_2[[val]], alternative = "two.sided", paired = FALSE, var.equal = TRUE)$p.value
      ## Wilcoxon test
      wilcoxon <- wilcox.test(data_group_1[[val]], data_group_2[[val]], alternative = "two.sided", paired = FALSE)$p.value
      
      data_results <- data.frame(group_1 = unique(data_group_1$group),
                                 group_2 = unique(data_group_2$group),
                                 shapiro_1 = shapiro_1, 
                                 shapiro_2 = shapiro_2,
                                 cohen_d = cohen_d,
                                 ttest_p = t,
                                 wilcoxon_p = wilcoxon)
      
      data_output <- bind_rows(data_output, data_results)
    }
  }
  return(data_output)
}
```

## Importing data
```{r}
data_raw <- read.csv(path_data, sep = ",")

head(data_raw)
```

## Normalising data
```{r}
data_normalised = data_raw %>%
  
  filter(treatment %in% c(0, 0.001953125, 0.0078125, 0.03125, 0.125, 0.5, 2)) %>%
  
  reframe(nuclei_count = n(),
          nuclei_area = sum(area_nucleus),
          DAPI_total = sum(integrated_int_DAPI),
          EdU_mean = mean(intensity_mean_EdU_nucleus) / mean(intensity_mean_EdU_cyto),
          p21_mean = mean(intensity_mean_p21_nucleus),
          .by = c(experiment, plate_id, well, well_id, image_id, cell_line, condition, siRNA, inhibitor, treatment, cyto_id, area_cell)) %>%
  
  group_by(experiment) %>%
  group_split() %>%
  lapply(fun_normalise, values = c("EdU_mean", "p21_mean"), ref = c("siCTR_MKI2_0")) %>%
  bind_rows() %>%
  mutate(EdU = case_when(EdU_mean_norm <= filter(data_thresholds, feature == "EdU")$T1 ~ "negative",
                         EdU_mean_norm > filter(data_thresholds, feature == "EdU")$T1 ~ "positive"),
         p21 = case_when(p21_mean_norm <= filter(data_thresholds, feature == "p21")$T1 ~ "negative",
                         p21_mean_norm > filter(data_thresholds, feature == "p21")$T1 ~ "positive")) %>%
  group_by(experiment, cell_line) %>%
  group_split() %>%
  lapply(fun_normalise, values = c("DAPI_total", "area_cell", "nuclei_area"), ref = c("siCTR_MKI2_0")) %>%
  bind_rows() %>%
  
  mutate(Hoechst = case_when(DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T1 ~ "Debris",
                             DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T2 ~ "2N",
                             DAPI_total_norm <= filter(data_thresholds, feature == "Hoechst")$T3 ~ "4N",
                             DAPI_total_norm > filter(data_thresholds, feature == "Hoechst")$T3 ~ "8N+")) %>%
  
  mutate(cell_cycle = case_when(Hoechst == "Debris" ~ "Debris",
                                Hoechst == "2N" & EdU == "negative" ~ "2N",
                                Hoechst == "4N" & EdU == "negative" ~ "4N",
                                Hoechst == "2N" & EdU == "positive" ~ "S",
                                Hoechst == "4N" & EdU == "positive" ~ "S",
                                Hoechst == "8N+" ~ "8N+",
                                .default = "Unassigned"))

head(data_normalised)
```

## Establishing cell counts
```{r}
data_cell_counts <- data_normalised %>%
  filter(cell_cycle != "Debris") %>%
  reframe(count = n(), .by = c(experiment, cell_line, siRNA, inhibitor, treatment)) %>%
  group_by(experiment, cell_line) %>%
  mutate(count_norm = count / max(count)) %>%
  ungroup()

data_cell_counts_summary = data_cell_counts %>%
  reframe(count_norm_mean = mean(count_norm),
          count_norm_sd = sd(count_norm),
          .by = c(cell_line, siRNA, inhibitor, treatment))

data_cell_counts
```

## Establishing proportions of cell cycle groups
```{r}
data_cell_cycle_proportions <- data_normalised %>%
  filter(cell_cycle != "Debris") %>%
  reframe(count = n(), .by = c(experiment, cell_line, siRNA, inhibitor, treatment, cell_cycle)) %>%
  group_by(experiment, cell_line, siRNA, inhibitor, treatment) %>%
  mutate(prop = 100 * count / sum(count)) %>%
  ungroup()

data_p21_proportions <- data_normalised %>%
  filter(!(cell_cycle %in% c("Debris", "S"))) %>%
  reframe(count = n(), .by = c(experiment, cell_line, siRNA, inhibitor, treatment, p21)) %>%
  group_by(experiment, cell_line, siRNA, inhibitor, treatment) %>%
  mutate(prop = 100 * count / sum(count)) %>%
  ungroup()

head(data_cell_cycle_proportions)
head(data_p21_proportions)
```

## Computing statistics for the cell cycle analysis
```{r}
data_stats_cell_cycle <- data_cell_cycle_proportions %>%
  filter(cell_cycle %in% c("4N", "8N+")) %>%
  complete(experiment, cell_line, siRNA, inhibitor, cell_cycle) %>%
  replace(is.na(.), 0) %>%
  group_by(cell_cycle) %>%
  group_split() %>%
  lapply((function (data) {
    
    tmp.cell.cycle <- unique(data$cell_cycle)
    statistics <- fun_stats(data, group_vars = c("cell_line", "siRNA", "inhibitor", "treatment"), val = "prop") %>%
      mutate(cell_cycle = tmp.cell.cycle)
    return(statistics)

  })) %>%
  bind_rows() %>%
  filter(group_1 != group_2)

data_stats_cell_cycle_subset <- data_stats_cell_cycle %>%
  separate(group_1, into = c("cell_line_1", "siRNA_1", "inhibitor_1", "treatment_1"), sep = "_", remove = FALSE) %>%
  separate(group_2, into = c("cell_line_2", "siRNA_2",  "inhibitor_2", "treatment_2"), sep = "_", remove = FALSE) %>%
  filter(cell_line_1 == cell_line_2, siRNA_1 == siRNA_2, inhibitor_1 == inhibitor_2, treatment_1 == 0, treatment_1 != treatment_2) %>%
  mutate(shapiro_1_adj = p.adjust(shapiro_1, method = "BH"),
         shapiro_2_adj = p.adjust(shapiro_2, method = "BH"),
         ttest_p_adj = p.adjust(ttest_p, method = "BH"),
         wilcoxon_p_adj = p.adjust(wilcoxon_p, method = "BH"))

data_stats_cell_cycle_subset
```

## Computing statistics for the proportion of p21-positive cells
```{r}
data_stats_p21 <- data_p21_proportions %>%
  filter(p21 == "positive") %>%
  complete(experiment, cell_line, siRNA, inhibitor, p21, treatment) %>%
  filter((inhibitor == "MKI-2" & treatment %in% c(0, 0.001953125, 0.0078125, 0.03125, 0.125, 0.5, 2)) |
         (inhibitor == "C-604" & treatment %in% c(0, 0.5, 2))) %>%
  replace(is.na(.), 0) %>%
  fun_stats(group_vars = c("cell_line", "siRNA", "inhibitor", "treatment"), val = "prop") %>%
  filter(group_1 != group_2)

data_stats_p21_subset <- data_stats_p21 %>%
  separate(group_1, into = c("cell_line_1", "siRNA_1", "inhibitor_1", "treatment_1"), sep = "_", remove = FALSE) %>%
  separate(group_2, into = c("cell_line_2", "siRNA_2",  "inhibitor_2", "treatment_2"), sep = "_", remove = FALSE) %>%
  filter(cell_line_1 == cell_line_2, siRNA_1 == siRNA_2, inhibitor_1 == inhibitor_2, treatment_1 == 0, treatment_1 != treatment_2) %>%
  mutate(shapiro_1_adj = p.adjust(shapiro_1, method = "BH"),
         shapiro_2_adj = p.adjust(shapiro_2, method = "BH"),
         ttest_p_adj = p.adjust(ttest_p, method = "BH"),
         wilcoxon_p_adj = p.adjust(wilcoxon_p, method = "BH"))

data_stats_p21_subset
```

## Plotting cell counts (heatmaps)
```{r}
list_cell_lines <- c("RPE-1", "HeLa")
list_siRNAs <- c("siCTR", "siMASTL", "siPPP2R2A", "siMASTLsiPPP2R2A")
list_treatments <- c("0", "0.001953125", "0.0078125", "0.03125", "0.125", "0.5", "2")

data_graph <- data_cell_counts %>%
  filter(cell_line %in% c(list_cell_lines)) %>%
  mutate(siRNA = factor(siRNA, levels = list_siRNAs, labels = c("siCTR", "siGWL", "siB55α", "siGWL\nsiB55α")),
         cell_line = factor(cell_line, levels = list_cell_lines),
         treatment = factor(treatment, levels = list_treatments, labels = c("0", "0.002", "0.008", "0.031", "0.125", "0.5", "2")),
         experiment = factor(experiment, levels = unique(experiment), labels = c(1, 2, 3)))

graph_cell_counts <- ggplot(data = data_graph,
                            aes(x = treatment,
                                y = interaction(experiment, cell_line),
                                fill = count_norm * 100,
                                group = interaction(siRNA, cell_line))) +
  
  geom_tile() +
  
  facet_nested(siRNA ~ inhibitor,
               nest_line = element_line(linetype = 1, linewidth = 0.25),
               resect = unit(0.04, "in"),
               strip = strip_nested(clip = "off"),
               scales = "free_x", space = "free") +
  
  scale_fill_gradient2(low = "#FFFFFF", mid = colour_pallette[4], high = colour_pallette[2],
                       midpoint = 50, limits = c(0, 100), breaks = c(0, 20, 40, 60, 80, 100),
                       na.value = colour_pallette[1]) +
  
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(expand = c(0, 0), guide = guide_axis_nested(delim = ".")) +
  labs(x = "Drug dose (µM)", fill = "Cell count (norm)") +
  
  graph_theme +
  theme(panel.spacing.y = unit(0.2, "lines"),
        panel.spacing.x = unit(0.2, "lines"),
        axis.title.y = element_blank(),
        ggh4x.axis.nestline = element_line(linewidth = 0.25, colour = "#000000"),
        axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"),
        axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), hjust = 0, vjust = 0.5, family = "Arial"),
        strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), vjust = 0, hjust = 0.7, family = "Arial"),
        legend.position = "top",
        legend.justification = "center",
        legend.key = element_blank(),
        legend.key.width = unit(10, 'pt'),
        legend.key.height = unit(4.5, 'pt'),
        legend.text = element_text(colour = "#000000", size = 6, margin = margin(t = 1)),
        legend.margin = margin(b = -10),
        legend.spacing.x = unit(1, "pt"),
        legend.title = element_text(colour = "#000000", size = 6, hjust = 0.5)) +
  
    guides(fill = guide_colorbar(label.position = "bottom",
                                 frame.colour = "#000000",
                                 frame.linewidth = 0.25,
                                 ticks.colour = "#000000",
                                 title.position = "top"))

ggsave(paste(path_results, "graph_cell_counts_heatmap.pdf", sep = ""), graph_cell_counts,
       width = 37.5, height = 70, units = "mm", device = cairo_pdf)
```

## Plotting proportions of 2N, S, 4N and 8N+ cell cycle groups
```{r}
list_cell_lines <- c("RPE-1", "HeLa")
list_siRNAs <- c("siCTR", "siMASTL", "siPPP2R2A", "siMASTLsiPPP2R2A")

data_graph <- data_cell_cycle_proportions %>%
  filter(cell_cycle != "debris", cell_line %in% c(list_cell_lines)) %>%
  complete(experiment, cell_line, nesting(inhibitor, siRNA, treatment, cell_cycle)) %>%
  replace(is.na(.), 0) %>%
  arrange(treatment, experiment, cell_cycle) %>%
  group_by(cell_line, inhibitor, siRNA, cell_cycle) %>%
  mutate(x_axis = 1:n()) %>%
  ungroup() %>%
  mutate(x_axis = case_when(inhibitor == "C-604" ~ x_axis - 9,
                            .default = x_axis),
         cell_cycle = factor(cell_cycle, levels = c("2N", "S", "4N", "8N+")),
         siRNA = factor(siRNA, levels = list_siRNAs, labels = c("siCTR", "siGWL", "siB55α", "siGWL\nsiB55α")),
         cell_line = factor(cell_line, levels = list_cell_lines))

data_stats <- data_stats_cell_cycle_subset %>%
  mutate(cell_line = cell_line_2, inhibitor = inhibitor_2, siRNA = siRNA_2, treatment = treatment_2, p = ttest_p) %>%
  arrange(treatment, cell_cycle) %>%
  group_by(cell_line, inhibitor, cell_cycle, siRNA) %>%
  mutate(significance = case_when(p < 0.001 ~ "***",
                                  p < 0.01 ~ "**",
                                  p < 0.05 ~ "*",
                                  .default = "ns"),
         x_axis = 1:n()) %>%
  filter(cell_cycle %in% c("4N", "8N+")) %>%
  group_by(cell_line, inhibitor, treatment, siRNA) %>%
  group_split() %>%
  lapply((function (data) {
    if(sum(data$significance == "ns") == 2) {
      return(head(data, 1))
    } else {
      return(filter(data, significance != "ns"))
    }
  })) %>%
  bind_rows() %>%
  arrange(treatment, cell_cycle) %>%
  group_by(cell_line, inhibitor, cell_cycle, siRNA) %>%
  mutate(x_axis = case_when(inhibitor == "C-604" ~ x_axis - 3, .default = x_axis),
         siRNA = factor(siRNA, levels = list_siRNAs, labels = c("siCTR", "siGWL", "siB55α", "siGWL\nsiB55α")),
         cell_line = factor(cell_line, levels = list_cell_lines))

graph_cell_cycle_proportions <- ggplot() +
  
  geom_bar(data = data_graph,
           aes(x = x_axis,
               y = prop,
               fill = cell_cycle,
               group = interaction(siRNA, treatment, cell_cycle)),
           stat = "identity", position = "stack",
           width = 0.95) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -8.385, xend = -8.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -5.5, xend = -5.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -2.5, xend = -2.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.25) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = 0.385, xend = 0.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +

  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 0.613, xend = 0.613, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +  
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 3.5, xend = 3.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 6.5, xend = 6.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +  
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 9.5, xend = 9.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 12.5, xend = 12.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 15.5, xend = 15.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 18.5, xend = 18.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 21.385, xend = 21.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +

  geom_hline(yintercept = c(0, 100), colour = colour_pallette[6], linetype = "solid", linewidth = 0.25) +
  
  geom_text(data = filter(data_stats, significance != "ns", cell_cycle == "4N", inhibitor == "C-604"),
            aes(x = x_axis * 3 + 2, y = 112, label = significance), size = 2.11, colour = colour_pallette[4], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance != "ns", cell_cycle == "8N+", inhibitor == "C-604"), 
            aes(x = x_axis * 3 + 2, y = 102, label = significance), size = 2.11, colour = colour_pallette[5], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance != "ns", cell_cycle == "4N", inhibitor == "MKI-2"),
            aes(x = x_axis * 3 + 2, y = 112, label = significance), size = 2.11, colour = colour_pallette[4], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance != "ns", cell_cycle == "8N+", inhibitor == "MKI-2"), 
            aes(x = x_axis * 3 + 2, y = 102, label = significance), size = 2.11, colour = colour_pallette[5], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance == "ns", inhibitor == "C-604"), 
            aes(x = x_axis * 3 + 2, y = 115, label = significance), size = 2.11, colour = "#666666", family = "Arial") +
  
  geom_text(data = filter(data_stats, significance == "ns", inhibitor == "C-604"), 
            aes(x = x_axis * 3 + 2, y = 115, label = significance), size = 2.11, colour = "#666666", family = "Arial") +
  
  geom_text(data = filter(data_stats, significance == "ns", inhibitor == "MKI-2"), 
            aes(x = x_axis * 3 + 2, y = 115, label = significance), size = 2.11, colour = "#666666", family = "Arial") +
  
  facet_nested(siRNA ~ cell_line + inhibitor,
               nest_line = element_line(linetype = 1, linewidth = 0.25),
               resect = unit(0.04, "in"),
               strip = strip_nested(clip = "off"),
               scales = "free_x", space = "free") +
  
  coord_cartesian(clip = "off") +
  labs(x = "Drug dose (µM)", y = "Proportion (%)") +
  
  scale_fill_manual(values = c(colour_pallette[2], colour_pallette[3], colour_pallette[4], colour_pallette[5])) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-7, -4 ,-1, 2, 5, 8, 11, 14, 17, 20),
                     labels = c(0, 0.5, 2, 0, 0.002, 0.008, 0.031, 0.125, 0.5, 2)) +
  
  
  scale_y_continuous(expand = c(0, 0), limits = c(-1.5, 122), breaks = c(0, 25, 50, 75, 100)) +
  graph_theme +
  theme(panel.border = element_blank(),
        panel.spacing.y = unit(0.05, "lines"),
        panel.spacing.x = unit(0.2, "lines"),
        axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"),
        strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), vjust = 0, hjust = 0.7, family = "Arial"),
        legend.position = "top",
        legend.justification = "left",
        legend.key = element_blank(),
        legend.key.width = unit(5.5, 'pt'),
        legend.key.height = unit(4, 'pt'),
        legend.spacing.x = unit(1.0, 'pt'),
        legend.text = element_text(colour = "#000000", size = 6, margin = margin(), family = "Arial"),
        legend.margin = margin(b = -10),
        legend.title = element_blank())
  
ggsave(paste(path_results, "graph_cell_cycle_proportions.pdf", sep = ""), graph_cell_cycle_proportions,
       width = 70, height = 60, units = "mm", device = cairo_pdf)
```

## Plotting proportions of p21-positive and p21-negative cells
```{r}
list_cell_lines <- c("RPE-1", "HeLa")
list_siRNAs <- c("siCTR", "siMASTL", "siPPP2R2A", "siMASTLsiPPP2R2A")

data_graph <- data_p21_proportions %>%
  filter(cell_line %in% c(list_cell_lines)) %>%
  
  complete(experiment, cell_line, nesting(inhibitor, siRNA, treatment, p21)) %>%
  replace(is.na(.), 0) %>%
  
  arrange(treatment, experiment, p21) %>%
  group_by(cell_line, inhibitor, siRNA, p21) %>%
  mutate(x_axis = 1:n()) %>%
  ungroup() %>%
  mutate(x_axis = case_when(inhibitor == "C-604" ~ x_axis - 9,
                            .default = x_axis),
         p21 = factor(p21, levels = c("negative", "positive"), labels = c("p21-negative", "p21-positive")),
         siRNA = factor(siRNA, levels = list_siRNAs, labels = c("siCTR", "siGWL", "siB55α", "siGWL\nsiB55α")),
         cell_line = factor(cell_line, levels = list_cell_lines))

data_stats <- data_stats_p21_subset %>%
  mutate(cell_line = cell_line_2, inhibitor = inhibitor_2, siRNA = siRNA_2, treatment = treatment_2, p = ttest_p) %>%
  arrange(treatment) %>%
  group_by(cell_line, inhibitor, siRNA) %>%
  mutate(significance = case_when(p < 0.001 ~ "***",
                                  p < 0.01 ~ "**",
                                  p < 0.05 ~ "*",
                                  .default = "ns"),
         x_axis = 1:n()) %>%
  bind_rows() %>%
  mutate(x_axis = case_when(inhibitor == "C-604" ~ x_axis - 3, .default = x_axis),
         siRNA = factor(siRNA, levels = list_siRNAs, labels = c("siCTR", "siGWL", "siB55α", "siGWL\nsiB55α")),
         cell_line = factor(cell_line, levels = list_cell_lines))

graph_p21_proportions <- ggplot() +
  
  geom_bar(data = data_graph,
           aes(x = x_axis,
               y = prop,
               fill = p21,
               group = interaction(siRNA, treatment, p21)),
           stat = "identity", position = "stack",
           width = 0.95) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -8.385, xend = -8.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -5.5, xend = -5.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = -2.5, xend = -2.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.25) +
  
  geom_segment(data = filter(data_graph, inhibitor == "C-604"), aes(x = 0.385, xend = 0.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +

  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 0.613, xend = 0.613, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +  
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 3.5, xend = 3.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 6.5, xend = 6.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +  
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 9.5, xend = 9.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 12.5, xend = 12.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 15.5, xend = 15.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 18.5, xend = 18.5, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +
  
  geom_segment(data = filter(data_graph, inhibitor == "MKI-2"), aes(x = 21.385, xend = 21.385, y = 0, yend = 100),
               colour = colour_pallette[6], linetype = "solid", linewidth = 0.2) +

  geom_hline(yintercept = c(0, 100), colour = colour_pallette[6], linetype = "solid", linewidth = 0.25) +
  
  geom_text(data = filter(data_stats, significance != "ns", inhibitor == "C-604"), 
            aes(x = x_axis * 3 + 2, y = 102, label = significance), size = 2.11, colour = colour_pallette[8], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance != "ns", inhibitor == "MKI-2"), 
            aes(x = x_axis * 3 + 2, y = 102, label = significance), size = 2.11, colour = colour_pallette[8], family = "Arial") +
  
  geom_text(data = filter(data_stats, significance == "ns", inhibitor == "C-604"), 
            aes(x = x_axis * 3 + 2, y = 115, label = significance), size = 2.11, colour = "#666666", family = "Arial") +
  
  geom_text(data = filter(data_stats, significance == "ns", inhibitor == "MKI-2"), 
            aes(x = x_axis * 3 + 2, y = 115, label = significance), size = 2.11, colour = "#666666", family = "Arial") +
  
  facet_nested(siRNA ~ cell_line + inhibitor,
               nest_line = element_line(linetype = 1, linewidth = 0.25),
               resect = unit(0.04, "in"),
               strip = strip_nested(clip = "off"),
               scales = "free_x", space = "free") +
  
  coord_cartesian(clip = "off") +
  labs(x = "Drug dose (µM)", y = "Proportion (%)") +
  
  scale_fill_manual(values = c(colour_pallette[7], colour_pallette[8])) +
  scale_x_continuous(expand = c(0, 0), breaks = c(-7, -4 ,-1, 2, 5, 8, 11, 14, 17, 20),
                     labels = c(0, 0.5, 2, 0, 0.002, 0.008, 0.031, 0.125, 0.5, 2)) +
  
  scale_y_continuous(expand = c(0, 0), limits = c(-1.5, 122), breaks = c(0, 25, 50, 75, 100)) +
  graph_theme +
  theme(panel.border = element_blank(),
        panel.spacing.y = unit(0.05, "lines"),
        panel.spacing.x = unit(0.2, "lines"),
        axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"),
        strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), vjust = 0, hjust = 0.7, family = "Arial"),
        legend.position = "top",
        legend.justification = "left",
        legend.key = element_blank(),
        legend.key.width = unit(5.5, 'pt'),
        legend.key.height = unit(4, 'pt'),
        legend.spacing.x = unit(1.0, 'pt'),
        legend.text = element_text(colour = "#000000", size = 6, margin = margin(), family = "Arial"),
        legend.margin = margin(b = -10),
        legend.title = element_blank())
  
ggsave(paste(path_results, "graph_p21_proportions.pdf", sep = ""), graph_p21_proportions,
       width = 70, height = 60, units = "mm", device = cairo_pdf)
```
