---
title: "TCGA"
output: html_document
date: "2024-08-15"
---

## Load packages 
```{r}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggpubr)
library(ggforce)
library(ggExtra)
library(ggtext)
library(ggfortify)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(ggsurvfit)
library(survminer)
library(survival)
library(readxl)
library(stringr)
library(scales)
```

## Establishing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1, family = "Arial"),
  panel.spacing.y = unit(0.1, "lines"),
  panel.spacing.x = unit(0.1, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.5),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2), vjust = 0, family = "Arial"),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2), family = "Arial"),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 3), family = "Arial"),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), family = "Arial"),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 6, margin = margin(r = 1), family = "Arial"),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000", "#ccdba2", "#889466", "#c6b2d1", "#654875")
```

## Set working directory
```{r}
path_data <- "/path/to/data/TCGA_processed_RNA-Seq_data.csv"
path_results <- "/path/to/results/"
```

## Importing data
```{r}
data_raw <- read.csv(path_data, sep = ",")

head(data_raw)
```

## Processing data
```{r warning=FALSE}

data_processed <- data_raw %>%
  mutate(sample_type = case_when(sample_type == "Primary Blood Derived Cancer - Peripheral Blood" ~ "Primary Tumor",
                                 sample_type == "Additional - New Primary" ~ "Primary Tumor",
                                 .default = sample_type)) %>%
  dplyr::filter(sample_type %in% c("Primary Tumor", "Solid Tissue Normal")) %>%
  mutate(sample_type = factor(sample_type, levels = c("Solid Tissue Normal", "Primary Tumor"), labels = c("N", "T"))) %>%
  group_by(project, sample_type) %>%
  mutate(expression = case_when(fpkm <= quantile(fpkm, 0.5) ~ "Low", fpkm > quantile(fpkm, 0.5) ~ "High")) %>%
  ungroup()

head(data_processed)
```

## Plotting cancer-specific expression profiles for GOI
```{r}
GOIs <- c("MASTL", "PPP2R2A")

for (GOI in GOIs) {
  
  data_graph <- data_processed %>% 
    filter(gene_name == GOI)
  
  data_graph_expression_profiles_summary <- data_graph %>%
    reframe(fpkm_median = median(fpkm),
            .by = c(project, gene_name, sample_type))
  
  data_expression_profiles_normality <- data_graph %>%
    group_by(project, gene_name, sample_type) %>%
    dplyr::filter(n() >= 3) %>%
    ungroup() %>%
    reframe(shapiro_p = shapiro.test(fpkm)$p.value, .by = c(project, gene_name, sample_type))
  
  data_expression_profiles_stats <- data_graph %>%
    dplyr::select(project, barcode, sample_type, gene_name, fpkm) %>%
    spread(key = sample_type, value = fpkm) %>%
    group_by(project, gene_name) %>%
    group_split() %>%
    lapply((function (data) {
      
      tmp.N <- data[["N"]][!(is.na(data[["N"]]))]
      tmp.T <- data[["T"]][!(is.na(data[["T"]]))]
      
      if (length(tmp.N) < 3 | length(tmp.T) < 3) {
        
        tmp.output <- data.frame(project = unique(data$project), p = NA, mean_diff = NA)

        } else {
  
        tmp.p <- wilcox.test(tmp.N, tmp.T, alternative = "two.sided", paired = FALSE)$p.value
        tmp.output <- data.frame(project = unique(data$project),
                                 gene_name = unique(data$gene_name),
                                 p = tmp.p,
                                 mean_diff = mean(tmp.T) - mean(tmp.N))
      }
      
      return(tmp.output)})) %>%
    
    bind_rows() %>%
    mutate(p_adj =  p.adjust(p, method = "BH"),
           significance = case_when(p_adj < 0.00001 ~ "*****",
                                    p_adj < 0.0001 ~ "****",
                                    p_adj < 0.001 ~ "***",
                                    p_adj < 0.01 ~ "**",
                                    p_adj < 0.05 ~ "*",
                                    p_adj >= 0.05 ~ "ns",
                                    is.na(p) ~ "nd"))
  
  graph_expression_profiles <- ggplot() +
    
    geom_quasirandom(data = data_graph, aes(x = sample_type, y = fpkm, colour = sample_type),
                     size = 0.75, alpha = 0.25, shape = 16) +
    geom_point(data = data_graph_expression_profiles_summary, aes(x = sample_type, y = fpkm_median),
               size = 0.75, stroke = 0.35, shape = 21, colour = "#000000", fill = "#FFFFFF") +
    
    geom_text(data = dplyr::filter(data_expression_profiles_stats, significance %in% c("nd", "ns")),
              aes(x = 1.5, y = 1.1 * max(data_graph$fpkm), label = significance),
              size = 6/.pt, colour = colour_pallette[1], family = "Arial") +
    
    geom_text(data = dplyr::filter(data_expression_profiles_stats, !(significance %in% c("nd", "ns")), mean_diff < 0),
              aes(x = 1.5, y = 1.1 * max(data_graph$fpkm), label = significance),
              size = 6/.pt, colour = colour_pallette[3], family = "Arial") +
    
    geom_text(data = dplyr::filter(data_expression_profiles_stats, !(significance %in% c("nd", "ns")), mean_diff > 0),
              aes(x = 1.5, y = 1.1 * max(data_graph$fpkm), label = significance),
              size = 6/.pt, colour = colour_pallette[2], family = "Arial") +
    
    scale_colour_manual(values = c(colour_pallette[3], colour_pallette[2])) +
    scale_y_continuous(limits = c(0, 1.2 * max(data_graph$fpkm))) +
    facet_wrap(~ project, nrow = 3) +
    labs(x = "Tissue (N - Normal, T - Tumor)", y = paste(GOI, " expression (FPKM)", sep = "")) +
    graph_theme
  
  ggsave(paste(path_results, GOI, "_graph_expression.pdf", sep = ""), graph_expression_profiles,
         width = 85, height = 50 , units = "mm", device = cairo_pdf)
}
```

## Processing the expression data
```{r}
model <- function(PPP2R2A, MASTL) {return(0.399 * MASTL / PPP2R2A + 0.406)}

data_expression_processed = data_raw %>%
  mutate(sample_type = case_when(sample_type == "Primary Blood Derived Cancer - Peripheral Blood" ~ "Primary Tumor", .default = sample_type)) %>%
  filter(sample_type %in% c("Primary Tumor", "Solid Tissue Normal")) %>%
  dplyr::select(barcode, project, gender, sample_type, gene_name, fpkm) %>%
  spread(key = gene_name, value = fpkm) %>%
  mutate(ED50_predicted = model(PPP2R2A, MASTL),
         sample_type = case_when(sample_type == "Primary Tumor" ~ "T", sample_type == "Solid Tissue Normal" ~ "N"),
         hit = ifelse(ED50_predicted < quantile(ED50_predicted, probs = 0.05), TRUE, FALSE))

data_expression_processed_hits <- data_expression_processed %>%
  filter(hit == TRUE) 

data_expression_processed
```

## Establishing counts of tissues predicted to be hyper sensitive to Greatwall inhibition
```{r}
data_sensitive_tissues_counts <- data_expression_processed %>%
  reframe(n = n(), .by = c(project, sample_type, hit)) %>%
  group_by(project, sample_type) %>%
  mutate(perc = n / sum(n)) %>%
  ungroup()

data_sensitive_tissues_counts
```

## Establishing significance of differences between tumour and normal tissues
```{r}
data_expression_profiles_stats <- data_expression_processed %>%
  dplyr::select(project, barcode, sample_type, MASTL, PPP2R2A) %>%
  gather(key = "gene", value = "fpkm", 4:5) %>%
  spread(key = sample_type, value = fpkm) %>%
  group_by(gene, project) %>%
  group_split() %>%
  lapply((function (data) {
    
    tmp.N <- data[["N"]][!(is.na(data[["N"]]))]
    tmp.T <- data[["T"]][!(is.na(data[["T"]]))]
    
    if (length(tmp.N) < 3 | length(tmp.T) < 3) {
      
      tmp.output <- data.frame(project = unique(data$project), p = NA)
    } else {
      
      tmp.p <- t.test(tmp.N, tmp.T, alternative = "two.sided", var.equal = FALSE)$p.value
      tmp.output <- data.frame(project = unique(data$project), gene = unique(data$gene), p = tmp.p, mean_diff = mean(tmp.T) - mean(tmp.N))
    }
    
    return(tmp.output)})) %>%
  
  bind_rows() %>%
  mutate(significance = case_when(p < 0.00001 ~ "*****",
                                  p < 0.0001 ~ "****",
                                  p < 0.001 ~ "***",
                                  p < 0.01 ~ "**",
                                  p < 0.05 ~ "*",
                                  p >= 0.05 ~ "ns",
                                  is.na(p) ~ "nd"))

data_expression_profiles_stats_processed <- data_expression_profiles_stats %>%
  filter(significance != "nd") %>%
  dplyr::select(project, gene, mean_diff) %>%
  spread(key = gene, value = mean_diff) %>%
  mutate(group = case_when(MASTL > 1 & PPP2R2A > 1 ~ "Group 1",
                           MASTL > 1 & PPP2R2A < 1 ~ "Group 2",
                           MASTL < 1 & PPP2R2A < 1 ~ "Group 3",
                           MASTL < 1 & PPP2R2A < 1 ~ "Group 4"))

data_expression_profiles_stats_processed
```

## Plotting mean FPKM differences between tumour and normal samples
```{r}
### Setting the working directory
setwd(working_directory)

data_graph <- data_expression_profiles_stats_processed
data_graph_counts <- reframe(data_graph, n = n(), .by = group)

graph_PPP2R2A_MASTL_mean_differences <- ggplot(data = data_graph, aes(x = MASTL, y = PPP2R2A, colour = group)) +
  geom_hline(yintercept = 1, linewidth = 0.25, colour = colour_pallette[6]) +
  geom_vline(xintercept = 1, linewidth = 0.25, colour = colour_pallette[6]) +
  geom_point(shape = 16, size = 1.25, alpha = 1) +
  
  annotate(geom = "text", x = -3.5, y = 5, label = 0, size = 6/.pt, family = "Arial", vjust = 0.9, hjust = 0.1, colour =  colour_pallette[2]) +
  annotate(geom = "text", x = 5, y = 5, label = 8, size = 6/.pt, family = "Arial", vjust = 0.9, hjust = 0.9, colour =  colour_pallette[3]) +
  annotate(geom = "text", x = -3.5, y = -3.5, label = 8, size = 6/.pt, family = "Arial", vjust = 0, hjust = 0.1, colour =  colour_pallette[6]) +
  annotate(geom = "text", x = 5, y = -3.5, label = 5, size = 6/.pt, family = "Arial", vjust = 0, hjust = 0.9, colour =  colour_pallette[5]) +
  
  scale_colour_manual(values = c(colour_pallette[3], colour_pallette[5], colour_pallette[1])) +
  scale_x_continuous(limits = c(-3.5, 5)) +
  scale_y_continuous(limits = c(-3.5, 5)) +
  labs(x = "GWL (T-N)", y = "B55α (T-N)") +
  
  graph_theme

ggsave(paste(getwd(), "/results/", "PPP2R2A_MASTL_mean_differences.pdf", sep = ""), graph_PPP2R2A_MASTL_mean_differences,
       width = 22.5, height = 25 , units = "mm", device = cairo_pdf)
```

## Plotting MASTL ~ PPP2R2A scatterplot
```{r warning=FALSE}
### Setting the working directory
setwd(working_directory)

data_graph <- data_expression_processed
data_graph_model <- data_expression_processed_hits

graph_scatterplot_MASTL_PPP2R2A <- ggplot() +
  
  geom_point(data = data_graph, aes(x = MASTL, y = PPP2R2A, colour = ED50_predicted),
             shape = 16, size = 0.35, alpha = 1) +
  
  geom_abline(slope = 5.656854249, colour = colour_pallette[6], linewidth = 0.25, linetype = "dashed") +
  
  annotate(geom = "text", x = 0, y = 54, label = "5%", size = 6/.pt, family = "Arial", hjust = 0.05, colour =  colour_pallette[6]) +
  
  scale_x_continuous(limits = c(0, 30), breaks = c(0, 15, 30)) +
  scale_y_continuous(limits = c(0, 55), breaks = c(0, 15, 30, 45)) +
  
  scale_colour_gradientn(colours = c(colour_pallette[2], colour_pallette[4], colour_pallette[7], colour_pallette[3], colour_pallette[5]),
                         values = seq(0, 1, 0.2),
                         limits = c(0.4, 2),
                         oob = oob_squish) +  

  labs(x = "GWL FPKM", y = "B55α FPKM", colour = bquote(Predicted~ED[50])) +

  graph_theme +
  theme(panel.border = element_blank(),
        plot.margin = margin(b = 0.2, l = 0.2, t = 0.35, r = 0.2, unit = "lines"),
        axis.line = element_line(colour = "#000000", linewidth = 0.25),
        legend.position = "right",
        legend.justification = "center",
        legend.key.width = unit(2.5, 'pt'),
        legend.key.height = unit(8, 'pt'),
        legend.text = element_text(colour = "#000000", size = 6, margin = margin(l = 2)),
        legend.margin = margin(l = -10),
        legend.spacing.x = unit(1, "pt"),
        legend.title = element_blank())

ggsave(paste(getwd(), "/results/", "scatterplot_MASTL_PPP2R2A.pdf", sep = ""), graph_scatterplot_MASTL_PPP2R2A,
       width = 25, height = 25 , units = "mm", device = cairo_pdf)
```

## Plotting distribution of ED50 predictions
```{r warning=FALSE}
### Setting the working directory
setwd(working_directory)

data_graph <- data_expression_processed %>%
  arrange(ED50_predicted) %>%
  mutate(rank = 1:n(), hit = ifelse(ED50_predicted < quantile(ED50_predicted, probs = 0.05), TRUE, FALSE))

graph_predicted_sensitivity_to_GWL_inhibition_distribution <- ggplot() +
  geom_point(data = data_graph, aes(x = rank, y = ED50_predicted, colour = hit),
             size = 0.1, position = position_jitter(0, 0.1), alpha = 1) +
  annotate(geom = "text", x = 100, y = 1, label = "5% most sensitive", family = "Arial", size = 6/.pt, hjust = 0, colour = colour_pallette[2]) +
  scale_x_continuous(expand = c(0.015, 0.015)) +
  scale_y_continuous(expand = c(0.03, 0.03), trans = "log2") +
  labs(x = "Rank", y = bquote(Model~"*ED"[50])) +

  scale_colour_manual(values = c(colour_pallette[1], colour_pallette[2])) +
  graph_theme +
  theme(panel.border = element_blank(),
        plot.margin = margin(b = 0.2, l = 0.2, t = 0.35, r = 0.2, unit = "lines"),
        axis.line = element_line(colour = "#000000", linewidth = 0.25))

ggsave(paste(getwd(), "/results/", "predicted_sensitivity_to_GWL_inhibition_distribution.pdf", sep = ""), graph_predicted_sensitivity_to_GWL_inhibition_distribution,
       width = 35, height = 20 , units = "mm", device = cairo_pdf)
```

## Plotting proportions of highly sensitive tissues
```{r warning=FALSE}
### Setting the working directory
setwd(working_directory)

data_graph <- data_sensitive_tissues_counts %>%
  mutate(perc = ifelse(sample_type == "N", - perc, perc))

data_graph_observations <- data_graph %>%
  reframe(n = sum(n), .by = c(project, sample_type)) %>%
  complete(project, sample_type) %>%
  replace(is.na(.), 0)

graph_proportions <- ggplot() +
  
  geom_bar(data = data_graph, aes(x = project, y = perc, fill = sample_type, alpha = hit),
           stat = "identity", position = "stack", width = 1) +
  
  geom_hline(yintercept = 0, colour = colour_pallette[6], linewidth = 0.25) +
  geom_vline(xintercept = seq(0.5, 32.5, 1), colour = colour_pallette[6], linewidth = 0.25) +
  
  geom_text(data = filter(data_graph_observations, sample_type == "N", n != 0),
            aes(x = project, y = -0.95, label = n),
            size = 5/.pt, colour = colour_pallette[3], family = "Arial",
            angle = -90, hjust = 1) +
  
  geom_text(data = filter(data_graph_observations, sample_type == "T", n != 0),
            aes(x = project, y = 0.95, label = n),
            size = 5/.pt, colour = colour_pallette[2], family = "Arial",
            angle = -90, hjust = 0) +
  
  geom_text(data = filter(data_graph_observations, sample_type == "N", n == 0),
            aes(x = project, y = -0.95, label = n),
            size = 5/.pt, colour = colour_pallette[6], family = "Arial",
            angle = -90, hjust = 1) +
  
  scale_fill_manual(values = c(colour_pallette[3], colour_pallette[2])) +
  scale_alpha_manual(values = c(0.25, 1), labels = c("Within remaining 95%", "Within 5% predicted most sensitive")) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0),
                     limits = c(-1, 1),
                     breaks = c(-1, -0.5, 0, 0.5, 1),
                     labels = c("100%", "50%", "0%", "50%", "100%"),
                     sec.axis = sec_axis(~.*1, name = "", breaks = c(-0.5, 0.5), labels = c("N", "T"))) +
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, hjust = 0, vjust = 0.5, family = "Arial"),
        axis.text.y.right = element_text(colour = "#000000", size = 6, margin = margin(l = 2), family = "Arial"),
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.y.right = element_blank(),
        legend.position = "top",
        legend.justification = "left",
        legend.key = element_blank(),
        legend.key.width = unit(5.5, 'pt'),
        legend.key.height = unit(5.5, 'pt'),
        legend.text = element_text(colour = "#000000", size = 6, margin = margin(l = 1)),
        legend.margin = margin(b = -10),
        legend.title = element_blank()) +
  
  guides(fill = FALSE)

ggsave(paste(getwd(), "/results/", "sensitive_groups_proportions.pdf", sep = ""), graph_proportions,
       width = 80, height = 26 , units = "mm", device = cairo_pdf)
```
