---
title: "Live Cell Imaging Analysis"
output: html_document
date: "2024-04-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(ggforce)
library(ggExtra)
library(grid)
library(gridExtra)
library(ggridges)
library(ggbeeswarm)
library(gghalves)
library(cowplot)
library(gplots)
library(lattice)
library(ggthemes)
library(scales)
library(seqinr)
library(tidyverse)
library(Dict)
library(viridis)
library(latex2exp)
library(ggnewscale)
library(ggh4x)
library(extrafont)
library(gtable)
library(rstatix)
library(readxl)
```

## Establishing paths to the data file and the folder to store results
```{r}
path_data <- "path/to/LC_cell_lineages.csv"
path_results <- "/path/to/results/"
```

## Establoshing ggplot theme arguments
```{r}
graph_theme <- theme(
  plot.margin = margin(b = 0.2, l = 0.2, t = 0.2, r = 0.2, unit = "lines"),
  text = element_text(size = 1, family = "Arial"),
  panel.spacing.y = unit(0.05, "lines"),
  panel.spacing.x = unit(0.05, "lines"),
  panel.border = element_rect(colour = "#000000", fill = NA, linewidth = 0.5),
  panel.background = element_blank(),
  panel.grid.major.y = element_blank(),
  panel.grid.minor.y = element_blank(),
  panel.grid.major.x = element_blank(),
  panel.grid.minor.x = element_blank(),
  strip.background = element_blank(),
  strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 2, t = 2), family = "Arial"),
  strip.text.y = element_text(colour = "#000000", size = 6, margin = margin(l = 2, b = 1), family = "Arial"),
  strip.clip = "off",
  axis.ticks.x = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.y = element_line(colour = "#000000", linewidth = 0.25),
  axis.ticks.length = unit(0.1, "lines"),
  axis.title.x = element_text(colour = "#000000", size = 6, margin = margin(t = 3), family = "Arial"),
  axis.title.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), family = "Arial"),
  axis.text.y = element_text(colour = "#000000", size = 6, margin = margin(r = 2), family = "Arial"),
  plot.title = element_blank(),
  legend.key = element_rect(fill = "#FFFFFF"),
  legend.position = "none",
  legend.justification = "left",
  legend.key.width = unit(2, 'pt'),
  legend.key.height = unit(2, 'pt'),
  legend.text = element_text(colour = "#000000", size = 6, margin = margin(r = 1), family = "Arial"),
  legend.margin = margin(r = 0, b = 0, l = 0, t = 0),
  legend.title = element_blank())
```

## Establishing colour pallette
```{r}
colour_pallette <- c("#D4D3CF", "#DC6B83", "#75B1CE", "#D8C367", "#526C94", "#000000", "#ccdba2", "#889466", "#c6b2d1", "#654875")
```

## Establishing lineages
```{r}
n_generations <- 10
labels <- seq(2 ^ n_generations, 65535, 2 ^ n_generations * 2)
df_lineages <- data.frame(lineage = labels) %>%
  mutate(low = lineage - (2 ** n_generations - 1),
         high = lineage + (2 ** n_generations - 1))

print(df_lineages)
```

## Importing data
```{r}
data_raw <- read.csv(path_data, sep = ",")
```

## Establishing a datframe containing time-stamps of mitotic and fusion events
```{r}
data_mitotic_events <- data_raw %>%
  filter(classification %in% c("object divided", "multipolar division", "object fusion", "object loss", "object left the frame")) %>%
  mutate(registered_objects = gsub("\\[|\\]", "", registered_objects)) %>%
  reframe(minimum = min(label), maximum = max(label), label = round(median(label)),
          .by = c(experiment, drug, cell_line, treatment, lineage, frame, time, classification, registered_objects)) %>%
  group_by(experiment, drug, cell_line, treatment, lineage, frame, time, classification, label) %>%
  group_split() %>%
  lapply((function (data) {
    
    if (data$classification %in% c("object divided", "multipolar division")) {
      
      objects <- strsplit(data$registered_objects, " ")[[1]]
      objects <- as.numeric(objects[objects != ""])
      data <- mutate(data, minimum = min(objects), maximum = max(objects))
    }
    return(data)
  })) %>%
  bind_rows() %>%
  ungroup()
 
data_mitotic_events 
```

## Establishing number of cell divisions and cell fusions per lineage
```{r}
data_divisions_per_lineage <- data_mitotic_events %>%
  filter(classification == "object divided") %>%
  reframe(count = n(), .by = c(experiment, cell_line, treatment, lineage)) %>%
  complete(experiment, cell_line, treatment, lineage) %>%
  replace(is.na(.), 0)

data_fusions_per_lineage <- data_mitotic_events %>%
  filter(classification == "object fusion") %>%
  reframe(count = n(), .by = c(experiment, cell_line, treatment, lineage)) %>%
  complete(experiment, cell_line, treatment, lineage) %>%
  replace(is.na(.), 0)

data_divisions_per_lineage
```

## Plotting cell lineages
```{r}
graph_data <- data_raw %>%
  filter(treatment %in% c(0, 0.25, 0.5, 1, 2), cell_line %in% c("HeLa", "RPE-1", "HCC1395")) %>%
  mutate(treatment = factor(treatment, levels = c(0, 0.25, 0.5, 1, 2)),
         cell_line = factor(cell_line, levels = c("HCC1395", "RPE-1", "HeLa")),
         drug = "C-604 (µM)")

graph_data_events <- data_mitotic_events %>%
  filter(treatment %in% c(0, 0.25, 0.5, 1, 2), cell_line %in% c("HeLa", "RPE-1", "HCC1395")) %>%
  gather(key = y, value = y_position, 10:11) %>%
  mutate(treatment = factor(treatment, levels = c(0, 0.25, 0.5, 1, 2)),
         cell_line = factor(cell_line, levels = c("HCC1395", "RPE-1", "HeLa")),
         drug = "C-604 (µM)")

graph_lineages <- ggplot() +
  
  geom_line(data = graph_data, aes(x = time / 60, y = label, group = label, colour = cell_line),
            linewidth = 0.5, alpha = 1, lineend = "round") +
  
  geom_line(data = graph_data_events, aes(x = time / 60, y = y_position, group = interaction(time, label), colour = cell_line),
            linewidth = 0.5, lineend = "round") +
  
  geom_point(data = filter(graph_data_events, classification == "object fusion"),
             aes(x = time / 60, y = label, fill = factor(lineage)),
             shape = 21, size = 0.75, stroke = 0.5, colour = colour_pallette[2], fill = "#FFFFFF") +
  
  geom_point(data = filter(graph_data_events, classification == "object loss"),
             aes(x = time / 60, y = label, fill = factor(lineage)),
             shape = 21, size = 0.75, stroke = 0.5, colour = "#000000", fill = "#000000") +
  
  scale_colour_manual(values = c(colour_pallette[1], colour_pallette[7], colour_pallette[9])) +
  
  scale_y_continuous(breaks = labels[1:10], labels = c(1:10)) +
  scale_x_continuous(breaks = c(0, 24, 48)) +
  labs(x = "Time (h)", y = "Lineage #") +
  facet_nested(. ~ cell_line + treatment, scales = "free",
               nest_line = element_line(linetype = 1, linewidth = 0.25),
               resect = unit(0.04, "in"),
               strip = strip_nested(clip = "off")) +
  graph_theme +
  theme(strip.text.x = element_text(colour = "#000000", size = 6, margin = margin(b = 1, t = 1), family = "Arial"),
        axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, vjust = 0.5, hjust = 0))

ggsave(paste(path_results, "graph_lineages.pdf", sep = ""), graph_lineages,
       width = 121.5, height = 48.5, units = "mm", device = cairo_pdf)
```

## Plotting number of cell divisions per lineage
```{r}
data_graph <- data_divisions_per_lineage %>%
  filter(treatment %in% c(0, 0.25, 0.5, 1, 2), cell_line %in% c("HeLa", "RPE-1", "HCC1395")) %>%
  mutate(treatment = factor(treatment, levels = c(0, 0.25, 0.5, 1, 2)),
         cell_line = factor(cell_line, levels = c("HCC1395", "RPE-1", "HeLa")),
         drug = "C-604 (µM)")

data_graph_summary <- data_graph %>%
  reframe(count_median = median(count), count_mean = mean(count), .by = c(experiment, cell_line, treatment))

graph_divisions_per_lineage <- ggplot() +
  
  geom_point(data = data_graph, aes(x = count, y = treatment, group = lineage, colour = cell_line),
             shape = 16, alpha = 0.75,
             size = 1, stroke = 0.25,
             position = position_jitter(0.21, 0.21)) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 2),
               aes(x = count_median, xend = count_median, y = 4.5, yend = 5.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 1),
               aes(x = count_median, xend = count_median, y = 3.5, yend = 4.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0.5),
               aes(x = count_median, xend = count_median, y = 2.5, yend = 3.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0.25),
               aes(x = count_median, xend = count_median, y = 1.5, yend = 2.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0),
               aes(x = count_median, xend = count_median, y = 0.5, yend = 1.5),
               colour = colour_pallette[2]) +
  
  scale_colour_manual(values = c(colour_pallette[1], colour_pallette[7], colour_pallette[9])) +
  geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5), colour = colour_pallette[1], linewidth = 0.25) +
  coord_flip() +
  labs(x = "Cell divisions per lineage", y = "C-604 (µM)") +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  facet_grid(cell_line ~ .) +
  
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, vjust = 0.5, hjust = 0))

ggsave(paste(path_results, "graph_divisions_per_lineage.pdf", sep = ""), graph_divisions_per_lineage,
       width = 22.5, height = 45, units = "mm", device = cairo_pdf)
```

## Plotting number of cell fusions per lineage
```{r}
data_graph <- data_fusions_per_lineage %>%
  filter(treatment %in% c(0, 0.25, 0.5, 1, 2), cell_line %in% c("HeLa", "RPE-1", "HCC1395")) %>%
  mutate(treatment = factor(treatment, levels = c(0, 0.25, 0.5, 1, 2)),
         cell_line = factor(cell_line, levels = c("HCC1395", "RPE-1", "HeLa")),
         drug = "C-604 (µM)")

data_graph_summary <- data_graph %>%
  reframe(count_median = median(count), count_mean = mean(count), .by = c(experiment, cell_line, treatment))

graph_fusions_per_lineage <- ggplot() +
  
  geom_point(data = data_graph, aes(x = count, y = treatment, group = lineage, colour = cell_line),
             shape = 16, alpha = 0.75,
             size = 1, stroke = 0.25,
             position = position_jitter(0.21, 0.21)) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 2),
               aes(x = count_median, xend = count_median, y = 4.5, yend = 5.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 1),
               aes(x = count_median, xend = count_median, y = 3.5, yend = 4.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0.5),
               aes(x = count_median, xend = count_median, y = 2.5, yend = 3.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0.25),
               aes(x = count_median, xend = count_median, y = 1.5, yend = 2.5),
               colour = colour_pallette[2]) +
  
  geom_segment(data = filter(data_graph_summary, treatment == 0),
               aes(x = count_median, xend = count_median, y = 0.5, yend = 1.5),
               colour = colour_pallette[2]) +
  
  scale_colour_manual(values = c(colour_pallette[1], colour_pallette[7], colour_pallette[9])) +
  geom_hline(yintercept = c(1.5, 2.5, 3.5, 4.5), colour = colour_pallette[1], linewidth = 0.25) +
  scale_fill_gradient2(low = "#FFFFFF", mid = colour_pallette[4], high = colour_pallette[2],
                       midpoint = 4, limits = c(0, 8), breaks = c(0),
                       na.value = colour_pallette[1]) +
  coord_flip() +
  labs(x = "Cell fusions per lineage", y = "C-604 (µM)") +
  scale_y_discrete(expand = c(0.1, 0.1)) +
  scale_x_continuous(expand = c(0.1, 0.1)) +
  facet_grid(cell_line ~ .) +
  
  graph_theme +
  theme(axis.text.x = element_text(colour = "#000000", size = 6, margin = margin(t = 2), angle = -90, vjust = 0.5, hjust = 0))

ggsave(paste(path_results, "graph_fusions_per_lineage.pdf", sep = ""), graph_fusions_per_lineage,
       width = 22.5, height = 45, units = "mm", device = cairo_pdf)
```
